---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# readjsa

<!-- badges: start -->
[![R-CMD-check](https://github.com/MattCowgill/readjsa/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/MattCowgill/readjsa/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

{readjsa} streamlines the process of getting data from [Jobs and Skills Australia](https://www.jobsandskills.gov.au) into R.

Note that this package is marked **experimental**. The code is stable, but brittle. Any changes by JSA to its data and how it's distributed are likely to break the functionality of the package, at least temporarily.

## Installation

You can install the development version of readjsa from [GitHub](https://github.com/) with :

``` r
# install.packages("devtools")
devtools::install_github("MattCowgill/readjsa")
```
## Usage

At the moment, `{readjsa}` can be used to get data from the Recruitment Experience & Outlook Survey (REOS), the Internet Vacancy Index (IVI), and Small Area Labour Markets (SALM). The functions for doing so are `read_reos`, `read_ivi`, and `read_salm` respectively - see below for more.

### Getting REOS data - `read_reos()`

It's straightforward to get data from the JSA REOS:

```{r reos-example}
library(readjsa)
library(ggplot2)
library(dplyr)

reos <- read_reos(tables = "all")

reos |>
  filter(
    series == "Recruitment rate",
    frequency == "Monthly"
  ) |>
  ggplot(aes(x = date, y = value, col = disaggregation)) +
  geom_line()
```

### Getting data from the JSA IVI - `read_ivi()`

There are several different tables in the IVI data, at different levels of aggregation. See `?read_ivi()`. In the example below, we're visualising total job vacancies at the state level. 

```{r ivi-example, eval=FALSE}
ivi_states <- read_ivi("2dig_states")

ivi_states |>
  filter(
    level == 1,
    state != "AUST"
  ) |>
  ggplot(aes(x = date, y = value, col = series_type)) +
  geom_line() +
  facet_wrap(~state,
    scales = "free_y",
    nrow = 2
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  )
```

### Getting SAML data - `read_salm()`

There are three SALM tables (total labour force in persons, total unemployment in persons, and unemployment rate) available by either Statistical Area Level 2 (SA2) and Local Government Area (LGA) levels. See `?read_salm()` for more information, and [the JSA website](https://www.jobsandskills.gov.au/data/small-area-labour-markets) for methodology and usage recommendations. 

In this example, we visualise smoothed estimates of unemployment rates across five LGAs in Perth since 2010.

```{r salm-example}
salm_LGA <- read_salm("unemp_rate")
perth_LGAs <- c("Stirling", "Joondalup", "Nedlands", "Swan", "Rockingham")

salm_LGA |>
  rename(LGA = local_government_area_lga_2023_asgs) |>
  filter(
    LGA %in% perth_LGAs
  ) |>
  ggplot(aes(
    x = quarter,
    y = value,
    col = LGA
  )) + 
  geom_line()
```

